Supplementary Materials for ``Frustration shapes multi-channel Kondo physics: a star graph perspective"

### 2023-01-09 manuscript/MCK-manuscript.tex
added orth cat.

\documentclass{iopart}
\usepackage{iopams}  
\usepackage{braket,graphicx,amsmath}
\usepackage[hidelinks]{hyperref}
\bibliographystyle{unsrt}
\begin{document}

@@

These polarised states can also be used to obtain the scattering phase shift suffered by the conduction channels at the Fermi surface. As shown in the Supplementary Materials~\cite{SM}, the impurity-bath hybdridisation in the polarised states contribute an excess charge of \(n_\mathrm{exc}^{(l)} = 1/K\) to each conduction channel. Following Friedel's sum rule, this leads to a total scattering phase shift of \(\pi/K\) for each channel:
\begin{eqnarray}
	\delta^{(l)} = \pi n_\mathrm{exc}^{(l)} = \frac{\pi}{K}, ~ ~ l \in \left[1, K\right] 
\end{eqnarray}
For \(K > 1\), each channel therefore experiences an unrenormalised phase shift of value between \(0\) and \(\pi\) that, as we will show in a later section, is responsible for the non-Fermi liquid behaviour. For example, we will show that this phase shift are linked to the critical non-Fermi liquid exponents of thermodynamic quantities at the fixed point. Right now, some evidence of this role of the phase shift can be provided through its relation to the vanishing field impurity magnetisation \(m\) obtained in eq.~\ref{imp-magnetisation}:
\begin{eqnarray}
	m(h=0^\pm) = \pm \frac{1}{2} \frac{\delta^{(l)} - \pi}{\delta^{(l)} + \pi}~.
\end{eqnarray}
When the phase shift \(\delta^{(l)}\) acquires a fractional value (in units of \(\pi\)), the impurity magnetisation also becomes fractional (in units of \(\frac{1}{2}\)). This shows that the unrenormalised phase shift value \(0 < \delta^{(l)} < \pi\) is at the heart of the quantum frustration present in the star graph (and therefore the MCK problem) that neither allows the impurity to be screened, nor allows it to decouple from the conduction channels. The decoupling only happens in the limit of extreme screening \((K \to \infty)\), and the impurity spin then becomes a local moment.

@@

@@

\bibliography{mscript_mck_full}

>>>>

\documentclass{iopart}
\usepackage{iopams}  
\usepackage{braket,graphicx,amsmath,color}
\usepackage[hidelinks]{hyperref}
\bibliographystyle{unsrt}
\begin{document}

@@

These polarised states can also be used to obtain the scattering phase shift suffered by the conduction channels at the Fermi surface. As shown in the Supplementary Materials~\cite{SM}, the impurity-bath hybdridisation in the polarised states contribute an excess charge of \(n_\mathrm{exc}^{(l)} = 1/K\) to each conduction channel. Following Friedel's sum rule, this leads to a total scattering phase shift of \(\pi/K\) for each channel:
\begin{eqnarray}
	\delta^{(l)} = \pi n_\mathrm{exc}^{(l)} = \frac{\pi}{K}, ~ ~ l \in \left[1, K\right] 
\end{eqnarray}
For \(K > 1\), each channel therefore experiences an unrenormalised phase shift of value between \(0\) and \(\pi\) that, as we will show in a later section, is responsible for the non-Fermi liquid behaviour. For example, we will show that this phase shift are linked to the critical non-Fermi liquid exponents of thermodynamic quantities at the fixed point. Right now, some evidence of this role of the phase shift can be provided through its relation to the vanishing field impurity magnetisation \(m\) obtained in eq.~\ref{imp-magnetisation}:
\begin{eqnarray}
	m(h=0^\pm) = \pm \frac{1}{2} \frac{\delta^{(l)} - \pi}{\delta^{(l)} + \pi}~.
\end{eqnarray}
When the phase shift \(\delta^{(l)}\) acquires a fractional value (in units of \(\pi\)), the impurity magnetisation also becomes fractional (in units of \(\frac{1}{2}\)). This shows that the "unrenormalised" phase shift value \(0 < \delta^{(l)} < \pi\)~\cite{anderson_1990} is at the heart of the quantum frustration present in the MCK problem, and leads, as we will see later, to a dramatic change in the ground-state wavefunction of the remaining degrees of freedom as the impurity is introduced, producing the well-known orthogonality catastrophe~\cite{anderson1967infrared,varma2002singular}.

@@

{\color{blue}We show in the Supplementary Materials~\cite{SM} that this orthogonality catastrophe can also be understood as a drastic change in the wavefunction and its quantum numbers. Upon taking account of the hopping between the star graph and the \(k-\)states of the conduction bath, the full ground-state \(\ket{\psi_-^{(l)}}\) and low-lying excited states \(\ket{\psi_+^{(l)}}\), in the \(l^\text{th}\) channel, take the form:
\begin{equation}\begin{aligned}
\ket{\psi_\pm^{(l)}} = \frac{1}{\sqrt 2}\left(\ket{\chi_{\frac{-1}{2}}}\otimes\ket{e_\uparrow^{(l)}} \pm \ket{\chi_{\frac{1}{2}}}\otimes\ket{e_\downarrow^{(l)}}\right)~,
\end{aligned}\end{equation}
where \(\ket{\chi_{\frac{\sigma}{2}}} = \frac{1}{\sqrt 6}\left( \ket{\sigma,\sigma,\bar\sigma} + \ket{\sigma,\bar\sigma,\sigma} -2\ket{\bar\sigma,\sigma,\sigma} \right),~ \sigma=\pm 1\) are the ground-states of the two-channel star graph, and \(\ket{e_\sigma^{(l)}} \equiv e^\dagger_{\sigma,(l)}\ket{\phi} = \sum_{k\in\text{FS}}c^\dagger_{k\sigma,(l)}\ket{\phi}\) are gapless excitations close to the Fermi surface (FS) \(\ket{\phi}\). We can now calculate the residue for scattering processes of the form \(e^\dagger_{\sigma,(l)} e_{\bar\sigma,(l)}\), in the new ground-state and excited state, and they are found to vanish:
\begin{equation}
	Z = |\braket{\psi_+^{(l)}| e^\dagger_{\uparrow,(l)} e_{\downarrow,(l)} | \psi_-^{(l)}}|^2 = \left|\bra{\psi_+^{(l)}}\left(\ket{\chi_{\frac{1}{2}}}\ket{e_\uparrow^{(l)}}\right)\right|^2 = 0
\end{equation}
This shows that the single-particle excitations are no longer long-lived, signaling the orthogonality catastrophe.}

@@

\bibliography{MCK-manuscript}

### 2023-01-09 manuscript/Supplementary/MCK-supplementary.tex
added orth cat.

\documentclass[10pt]{iopart}
\usepackage{iopams}  
\usepackage{braket,graphicx,amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=2cm]{geometry}
\bibliographystyle{unsrt}
\begin{document}

@@

\subsection{Orthogonality catastrophe in the gapless excitations}

\subsection{Three channel LEH}
Similar to the two channel case, we here calculate the low energy effective Hamiltonian for three channel Kondo problem by introducing the real space hopping on top of the zero mode three-channel star graph model. This zero mode three channel star graph model has three fold degenerate ground states with total $2^4=16$ states in the eigenspectrum. The three degenerate ground states are given as 
\begin{eqnarray}
|\alpha_{-1}\rangle = c|1000 \rangle-b (|0100 \rangle + |0010 \rangle+ | 0001 \rangle), ~ ~ |\alpha_{+1}\rangle = b(|1110 \rangle+ | 1101 \rangle + | 1011 \rangle)-c | 0111 \rangle,\nonumber\\
|\alpha_0\rangle = -a(|1100 \rangle + |1010\rangle +|1001 \rangle) + a(|0110 \rangle+| 0101 \rangle +| 0011\rangle)~,
\end{eqnarray}
where $a=0.408$, $b=0.289$, $c=0.866$. Here the state is represented by $|n_{d},n_1,n_2,n_3\rangle$, where $n_i=1/0$ represents the spin configuration $S_i^z=\pm 1/2$ respectively. Next we use degenerate perturbation theory to get the LEH which contains diagonal and off-diagonal terms. In this case we get non-zero contribution from all the three ground states $|J^z=-1\rangle$, $|J^z=0\rangle$ and $|J^z=1\rangle$. We get the diagonal contribution to the LEH in the second order to be
\begin{eqnarray}
H^{(2)}_{eff, diag} &=& - \frac{7.2 J^2}{\alpha} \hat{I}~.
\end{eqnarray}
The contribution associated with different ground states $|J^z=-1\rangle$, $|J^z=0\rangle$ and $|J^z=1\rangle$ is given respectively as $- \frac{2.4 J^2}{\alpha} \hat{I} + \hat{\mathcal{F}},- \frac{2.4 J^2}{\alpha} \hat{I} ,- \frac{2.4 J^2}{\alpha} \hat{I} - \hat{\mathcal{F} }$,
where $\hat{\mathcal{F}}$ is a function of diagonal number operators of sites nearest neighbor to the zeroth site of each channels. The off-diagonal terms in the LEH is appearing due to the scattering between pair of degenerate states $(\alpha_0,\alpha_{+1})$ and $(\alpha_0,\alpha_{-1})$, there is no contribution in the second order coming from the scattering between $(\alpha_{-1},\alpha_{+1})$. The effective low energy Hamiltonian in the second order is given as
\begin{eqnarray}
	H^{(2)}_{eff,off-diag} = \sum_{\substack{(ijk)=\\(123),(231),(312)}}&\left[ c_{i\uparrow}c_{i\downarrow}^{\dagger} \left( -2ab \left\{ \Sigma_{jk} +c_{j\uparrow}^{\dagger}c_{j\downarrow}c_{k\uparrow}c_{k\downarrow}^{\downarrow} \right\} -ac(\Omega_{jk}+\tilde{\Omega}_{jk}) \right) \right.\nonumber\\
										  &\left. + \textrm{h.c.}\right] \otimes \hat{\Xi}_l~,
\end{eqnarray}
where the different operators have the following definitions:
\begin{eqnarray}
\fl \Sigma_{i,j} = n_{i\uparrow}(1-n_{i\downarrow}) (1-n_{j\uparrow})n_{j\downarrow} + (1-n_{i\uparrow})n_{i\downarrow} n_{j\uparrow} (1 -n_{j\downarrow}), ~ ~\Omega_{i,j} =4S_i^z S_j^z n_{i\uparrow} n_{j\uparrow}, ~ ~ \tilde{\Omega}_{i,j} = 4S_{i}^zS_j^z (1-n_{i\uparrow})(1-n_{j\uparrow}) \nonumber~,\\
\fl \hat{\Xi}_l=  ( c_{l_1\uparrow}^{\dagger} c_{l_1\downarrow} +c_{l_2\uparrow}^{\dagger} c_{l_2\downarrow} + c_{l_3\uparrow}^{\dagger} c_{l_3\downarrow} + \textrm{h.c.})~.
\end{eqnarray}

@@

\bibliography{../mscript_mck_full}
\end{document}

>>>>

\documentclass[12pt]{iopart}
\usepackage{iopams}  
\usepackage{braket,graphicx,amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=2cm]{geometry}
\bibliographystyle{unsrt}
\begin{document}

@@

\subsection{Orthogonality catastrophe in the gapless excitations of the 2CK}
By combining the ground-states of the star graph and the effect of hopping into the remaining lattice, it is possible to demonstrate an orthogonality catastrophe~\cite{anderson1967infrared,varma2002singular} in the new modified ground-state. The full Hilbert space consists of the star graph (formed by the impurity site and the zeroth sites of all channels) and the {\it reduced} conduction bath (formed by sites 1 and beyond in all channels). All terms in the Hamiltonian conserve the total spin \(S_\text{tot}^z = S_d^z + S_0^z + s^z\) composed of the impurity spin \(S_d^z\), the zeroth site spin \(S_0^z = \sum_l S_{0,(l)}^z\) and the spin of the reduced bath \(s^z\). We will work in the \(S_\text{tot}^z = 0\) sector. To construct the non-magnetic ground-states, we will assume the presence of gapless excitations vanishingly close to the Fermi surface in the reduced conduction bath. If \(\ket{\phi}\) represents the filled Fermi volume configuration of all the channels, these excitations can be written as \(\ket{e_\sigma^{(l)}} \equiv e^\dagger_{\sigma,(l)} \ket{\phi} = \sum_{k \in \text{FS}}c^\dagger_{k\sigma,(l)} \ket{\phi}\), where \(l \in \left\{1,2\right\} \) is the channel index. Apart from these states, we have the star graph ground-states at \(S_\text{sg}^z \equiv S_d^z + S_0^z = \pm \frac{1}{2}\): \(\ket{S_\text{sg}^z = \frac{\sigma}{2}} = \frac{1}{\sqrt 6}\left( \ket{\sigma,\sigma,\bar\sigma} + \ket{\sigma,\bar\sigma,\sigma} -2\ket{\bar\sigma,\sigma,\sigma} \right),~ \sigma=\pm 1 \). Within each ket, the three symbols indicate the configuration of the impurity spin, the zeroth site of the first channel and that of the second channel respectively. In the absence of any hopping between the star graph and the reduced conduction bath, the full non-magnetic ground states are then of the form:
\begin{equation}\begin{aligned}
	\ket{l,\sigma} = \ket{S^z_\text{sg}=\frac{-\sigma}{2}}\otimes\ket{e_{\sigma}^{(l)}},~l\in\left\{ 1,2 \right\},~\sigma=\pm 1~.
\end{aligned}\end{equation}
We now introduce the hopping \(H_\text{int}\) between the star graph and the reduced conduction bath,
\begin{equation}\begin{aligned}
	H_\text{int} = -t\frac{1}{\sqrt N}\sum_{k,\sigma,l}\left( c^\dagger_{k\sigma,(l)}c_{0\sigma,(l)} + \text{h.c.} \right)~,
\end{aligned}\end{equation}
and treat it under degenerate perturbation theory. Since all terms in the Hamiltonian conserve the number of particles in each channel, the states with different \(l\) do not mix. For each value of \(l\), the \(2\times 2\) matrix of \(H_\text{int}\) takes the form
\begin{equation}\begin{aligned}
	H_\text{int} = \frac{4t^2}{J^*}\begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix} ~.
\end{aligned}\end{equation}
The non-zero off-diagonal terms arise because of second-order scattering processes.
Diagonalising this gives the modified new ground-states:
\begin{equation}\begin{aligned}
	\ket{\psi_\pm^{(l)}} = \frac{1}{\sqrt 2}\left(\ket{l,\uparrow} \pm \ket{l,\downarrow}\right), E_\pm = -4t^2/J^*~.
\end{aligned}\end{equation}
We can now calculate the residue for scattering processes of the kind \(e^\dagger_{\sigma,(l)} e_{\bar\sigma,(l)}\), in the new ground-state and excited state, and they are found to vanish:
\begin{equation}\begin{aligned}
	Z = |\braket{\psi_+^{(l)}| e^\dagger_{\uparrow,(l)} e_{\downarrow,(l)} | \psi_-^{(l)}}|^2 = \left|\bra{\psi_+^{(l)}}\left(\ket{S_\text{sg}^z=\frac{1}{2}}\ket{e_\uparrow^{(l)}}\right)\right|^2 = 0
\end{aligned}\end{equation}
This shows that the single-particle excitations are no longer long-lived, signaling the orthogonality catastrophe.

\subsection{Three channel LEH}
Similar to the two channel case, we here calculate the low energy effective Hamiltonian for three channel Kondo problem by introducing the real space hopping on top of the zero mode three-channel star graph model. This zero mode three channel star graph model has three fold degenerate ground states with total $2^4=16$ states in the eigenspectrum. The three degenerate ground states are given as 
\begin{eqnarray}
\fl |\alpha_{-1}\rangle = c|1000 \rangle-b (|0100 \rangle + |0010 \rangle+ | 0001 \rangle), ~ ~ |\alpha_{+1}\rangle = b(|1110 \rangle+ | 1101 \rangle + | 1011 \rangle)-c | 0111 \rangle,\nonumber\\
\fl |\alpha_0\rangle = -a(|1100 \rangle + |1010\rangle +|1001 \rangle) + a(|0110 \rangle+| 0101 \rangle +| 0011\rangle)~,
\end{eqnarray}
where $a=0.408$, $b=0.289$, $c=0.866$. Here the state is represented by $|n_{d},n_1,n_2,n_3\rangle$, where $n_i=1/0$ represents the spin configuration $S_i^z=\pm 1/2$ respectively. Next we use degenerate perturbation theory to get the LEH which contains diagonal and off-diagonal terms. In this case we get non-zero contribution from all the three ground states $|J^z=-1\rangle$, $|J^z=0\rangle$ and $|J^z=1\rangle$. We get the diagonal contribution to the LEH in the second order to be
\begin{eqnarray}
H^{(2)}_{eff, diag} &=& - \frac{7.2 J^2}{\alpha} \hat{I}~.
\end{eqnarray}
The contribution associated with different ground states $|J^z=-1\rangle$, $|J^z=0\rangle$ and $|J^z=1\rangle$ is given respectively as $- \frac{2.4 J^2}{\alpha} \hat{I} + \hat{\mathcal{F}},- \frac{2.4 J^2}{\alpha} \hat{I} ,- \frac{2.4 J^2}{\alpha} \hat{I} - \hat{\mathcal{F} }$,
where $\hat{\mathcal{F}}$ is a function of diagonal number operators of sites nearest neighbor to the zeroth site of each channels. The off-diagonal terms in the LEH is appearing due to the scattering between pair of degenerate states $(\alpha_0,\alpha_{+1})$ and $(\alpha_0,\alpha_{-1})$, there is no contribution in the second order coming from the scattering between $(\alpha_{-1},\alpha_{+1})$. The effective low energy Hamiltonian in the second order is given as
\begin{eqnarray}
	H^{(2)}_{eff,off-diag} = \sum_{\substack{(ijk)=\\(123),(231),(312)}}&\left[ c_{i\uparrow}c_{i\downarrow}^{\dagger} \left( -2ab \left\{ \Sigma_{jk} +c_{j\uparrow}^{\dagger}c_{j\downarrow}c_{k\uparrow}c_{k\downarrow}^{\downarrow} \right\} -ac(\Omega_{jk}+\tilde{\Omega}_{jk}) \right) \right.\nonumber\\
										  &\left. + \textrm{h.c.}\right] \otimes \hat{\Xi}_l~,
\end{eqnarray}
where the different operators have the following definitions:
\begin{eqnarray}
\fl \Sigma_{i,j} = n_{i\uparrow}(1-n_{i\downarrow}) (1-n_{j\uparrow})n_{j\downarrow} + (1-n_{i\uparrow})n_{i\downarrow} n_{j\uparrow} (1 -n_{j\downarrow}), ~ ~\Omega_{i,j} =4S_i^z S_j^z n_{i\uparrow} n_{j\uparrow}, ~ ~ \tilde{\Omega}_{i,j} = 4S_{i}^zS_j^z (1-n_{i\uparrow})(1-n_{j\uparrow}) \nonumber~,\\
\fl \hat{\Xi}_l=  ( c_{l_1\uparrow}^{\dagger} c_{l_1\downarrow} +c_{l_2\uparrow}^{\dagger} c_{l_2\downarrow} + c_{l_3\uparrow}^{\dagger} c_{l_3\downarrow} + \textrm{h.c.})~.
\end{eqnarray}

@@

\bibliography{../MCK-manuscript}
\end{document}
